\quad Massive props must go out to David Craddock and the team at Storybundle for providing me with the opportunity and motivation to create this book. It’s wonderful to have a platform for this niche content and I am hoping that Storybundle continues to provide opportunities for those dedicated to history and preservation to find an audience.

My main inspirations in this field have been Keith Smith and Alex Smith (no relation). Without them, I would not have taken the step in interviewing as many people as I have. Their dedicated research has opened up new doors for anyone getting involved with video game history and I could not be where I am in my journey without them.

The supportive crew at Gaming Alexandria has also provided me with opportunities to help tell stories from all across the industry and generally make the preservation community more vibrant. Special regards to Dustin Hubbard, Weiste van Bruggen, DillyDylan, Jonas Rosland, and Stefan Gancer for their continued efforts.

Of course there’s also Kate Willært to thank for the cover. We worked together to find something which spoke to the era and carried a sleekness of modern design. The color pallet was chosen based on the flyer for the Atari arcade game Crash 'n Score, which struck me as entirely unique yet also fitting in the 70s. Kate had much fun

The interviews were recorded using Audacity and Reaper and transcribed using Youtube’s auto-caption system as well as Descript. LibreOffice and Google Docs both provided the writing backbone, with the Overleaf LaTeX system allowing me to compile the words for easy reading. MAME and the Internet Archive allowed me to provide game screenshots to include as well. Thanks to all the supportive software developers that provide solutions both open source and commercial to help make things like Ebooks a reality.

Then of course to the interviewees, not only the ones I have included here but the over 100 people I’ve interviewed which have been part of the video game industry since it’s foundation. Thank you for making this history so vibrant and accessible; I hope to do your accomplishments justice in telling these stories.